---
title: "NCAA Basketball Coaches Salary Data 2013"
author: "Sumedh Deepak Patkar"
date: "2023-02-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

### Summary 🏀

The National Collegiate Athletic Association (NCAA) published in 2013 the Salary data of Men's basketball team coaches. The excel sheet can be found on this [**link**](https://github.com/Sumedh-Patkar/ST516/blob/main/data/1_2013_NCAA_basketball_coaches_salaries.xls?raw=true). 

The sheet contains salaries paid to the coaches of 62 teams in a recent NCAA men's Division 1 basketball tournament. We consider these 62 salaries as *sample data* from the population of 346 NCAA Division 1 Men's basketball coaches in the year the tournament was held.

In this RMarkdown sheet, we will load, visualize, summarize the data. We will then calculate a **98% confidence interval** for the coach salaries of the population.

---

### **Loading the data 📁**


```{r message = FALSE}
library("dplyr")
library("readxl")
library("ggplot2")
library("scales")

# Set the working directory to code directory
# You may need to change this variable if you try running this code on your machine
current_working_dir = "~/Development/ST516/1.NC_AA_Basketball_Coaches_Salaries"
setwd(current_working_dir)

# Read the excel file using readxl package function 'read_excel()'
salaries_xl <- readxl::read_excel("../data/1_2013_NCAA_basketball_coaches_salaries.xls", sheet = 1)

salaries_xl <- salaries_xl %>%
  rename(athletic_conference = CONF, school=SCHOOL, head_coach = `HEAD COACH`, school_pay = `SCHOOL PAY`,
         other_pay = `OTHER PAY`, total_pay = `TOTAL PAY`, max_bonus = `MAX BONUS`)
```

---

### **Visualizing the data 📈 📊 📉**

#### **Tabular view (First 6 rows only)**

```{r message=FALSE}
#Tabular representation of data
knitr::kable(head(salaries_xl))

```

#### **Summary of the data**

```{r message=FALSE}
# We skip the categorical attributes here. We will summarize some of them graphically below.
salaries_xl %>%
  dplyr::select(-'RANK', -'head_coach',-'school', -'athletic_conference') %>%
  summary() %>% 
  knitr::kable()
```

### **Graphical summary views**
##### **1. Representation of teams from various athletic conferences**


Following chart shows the number of teams belonging to various [athletic conferences](https://en.wikipedia.org/wiki/List_of_NCAA_conferences)

```{r message=FALSE}

ggplot(data = salaries_xl) + 
  geom_bar(mapping = aes(y = athletic_conference))


```

##### **2. Histogram of Total Pay**

Following chart shows the histogram of the total pay column

```{r message=FALSE}

ggplot(data = salaries_xl) +
  geom_histogram(mapping = aes(x = total_pay), colour="black", fill="white") + 
  scale_x_continuous(labels = comma)

```

Notice the outlier in the Total Pay.
